### 2015-09-12 docs/prelim_analysis.Rmd
updated prelim ms with new results

---
output:
  pdf_document:
    fig_caption: true
    keep_tex: true
fontsize: 12pt
geometry: margin=1in
---
# Drivers of species synchrony in real communities
*Andrew Tredennick, Peter Adler, Claire de Mazancourt, Michele Loreau, and Stephen Ellner*

Quick summary
-------------
1. __Question__. How (a)synchronous are species population dynamics in natural communities, and what determines the degree of synchrony among species? Of particular interest to us, do species' responses to environmental fluction increase or decrease population synchrony?

2. __Data__. Five longterm demographic datasets from semi-arid North American grasslands. All species coexist stabley, yet are similar -- they are all perennial grasses.

3. __Analysis__. (a) Calculate community synchrony from the observed time series using an established metric, and compare this to expectation from theory. (b) Use IPMs with environmental or demographic stochasticity turned on/off to quantify the relative contribution of each to community synchrony.

4. __Initial conclusions__. Species' responses to environmental fluctuations tend to synchronize population dynamics: at four of the five sites removing environmental fluctuations decreased community synchrony. However, even with environmental fluctuations present, populations were not completely synchronized. This implies that niche differences among these species remain strong even as they respond similarly to environmental conditions.

@@

```{r read_data, echo=FALSE}
output_list <- readRDS("../results/ipm_simulation_lists.RDS")
mlist <- melt(output_list)
```

```{r get_synchs, echo=FALSE, fig.cap="Results from IPM simulations with difference sources of stochasticity present/absent at each site. DEMO = demographic stochasticity only; ENV = environmental stochasticity only; ENVDEMO = both demograhpic and environmental stochasticity.", fig.width=6}
colnames(mlist)[1:3] <- c("year", "species", "cover")
mlist <- subset(mlist, year>500)
sites <- unique(mlist$L1)
sims <- unique(mlist$L2)
pgr_list <- list()
for(dosite in sites){
  tmp_data <- subset(mlist, L1==dosite)
  if(dosite=="Idaho"){
    tmp_data <- subset(tmp_data, species!="ARTR")
  }
  within_site_list <- list()
  for(dosim in sims){
    tmpsim <- subset(tmp_data, L2==dosim)[c("year", "species", "cover")]
    tmpsynch <- get_ipm_synchrony(tmpsim)
    tmp_pgr_synch <- as.numeric(tmpsynch$pgr_synchrony["obs"])
    within_site_list[[dosim]] <- tmp_pgr_synch
  }
  pgr_list[[dosite]] <- within_site_list
}
pgr_synch <- melt(pgr_list)

ggplot(pgr_synch, aes(x=L2, y=value, fill=L2))+
  geom_bar(stat="identity", position=position_dodge(0.9), color="grey")+
  facet_wrap("L1", scales="free_x")+
  scale_fill_manual(values = c("grey45", "steelblue", "slateblue4"))+
  guides(fill=FALSE)+
  scale_y_continuous(limits=c(0,1))+
  xlab("Source of stochasticity")+
  ylab("Species synchrony")

@@

```{r niches, echo=FALSE}
# niche_diff <- data.frame(site=sites,
#                          ND = c(0.58, 0.59, 0.66, 0.69, 0.68))
# pgr_subs <- subset(pgr_synch, L2!="ENVDEMO")
# pgr_cast <- dcast(pgr_subs, L1~L2, value.var = "value")
# synch_diff <- ddply(pgr_cast, .(L1), summarise,
#                     value = ENV-DEMO)
# niche_synch <- merge(niche_diff, synch_diff, by.x="site", by.y="L1")
# ggplot(niche_synch, aes(x=ND, y=value))+
#   geom_point(size=5)+
#   ylab("Community synchrony")
```

>>>>

---
layout: 12pt
header-includes:
   - \usepackage{lineno}
   - \linenumbers
   - \usepackage{setspace}
   - \doublespacing
   - \usepackage{todonotes}
   - \usepackage[document]{ragged2e}

## rmarkdown render options
output:
  pdf_document:
    fig_caption: true
    keep_tex: true
fontsize: 12pt
geometry: margin=1in

---
# Environmental responses synchronize population \newline dynamics in five semi-arid grasslands

\renewcommand*{\thefootnote}{\fnsymbol{footnote}}

Andrew T. Tredennick\footnote{Correspondance: atredenn@gmail.com}\textsuperscript{1}, Claire de Mazancourt\textsuperscript{2}, Michel Loreau\textsuperscript{2}, \newline Stephen Ellner\textsuperscript{3}, and Peter B. Adler\textsuperscript{1}

\textit{\small{\textsuperscript{1}Department of Wildland Resources and the Ecology Center, 5230 Old Main Hill, Utah State University, Logan, Utah 84322 USA}}

\textit{\small{\textsuperscript{2}Centre for Biodiversity Theory and Modelling, Experimental Ecology Station, Centre National de la Recherche Scientifique, Moulis, 09200, France}}

\textit{\small{\textsuperscript{3}Department of Ecology and Evolutionary Biology, Cornell University, Ithaca, NY 14853, USA}}

\renewcommand*{\thefootnote}{\arabic{footnote}}
\setcounter{footnote}{0}

Abstract
-------------
Temporal asynchrony among interacting species is an important mechanism through which species richness can stabilize biomass production through time. Theory

__Keywords__
synchrony, compensatory dynamics, environmental stochasticity, demographic stochasticity, stability

@@

```{r read_data, echo=FALSE}
output_list <- readRDS("../results/ipm_comp_nocomp_sims.RDS")
mlist <- melt(output_list)
```

```{r get_synchs, echo=FALSE, fig.cap="Results from IPM simulations with difference sources of stochasticity present/absent at each site. DEMO = demographic stochasticity only; ENV = environmental stochasticity only; ENVDEMO = both demograhpic and environmental stochasticity.", fig.width=6}
colnames(mlist)[1:3] <- c("year", "species", "cover")
sites <- unique(mlist$L1)
sims <- unique(mlist$L2)
pgr_list <- list()
for(dosite in sites){
  tmp_data <- subset(mlist, L1==dosite)
  if(dosite=="Idaho"){
    tmp_data <- subset(tmp_data, species!="ARTR")
  }
  within_site_list <- list()
  for(dosim in sims){
    tmpsim <- subset(tmp_data, L2==dosim)[c("year", "species", "cover")]
    tmpsynch <- get_ipm_synchrony(tmpsim)
    tmp_pgr_synch <- as.numeric(tmpsynch$pgr_synchrony["obs"])
    within_site_list[[dosim]] <- tmp_pgr_synch
  }
  pgr_list[[dosite]] <- within_site_list
}
pgr_synch <- melt(pgr_list)
expected_synch <- table_data[,c(1,3)]
expected_synch$L2 <- "E(SYNCH)"
colnames(expected_synch) <- c("L1", "value", "L2")
expected_synch <- expected_synch[,c("value", "L2", "L1")]
expected_synch[which(expected_synch[,"L1"]=="New Mexico"),"L1"] <- "NewMexico"

pgr_synch <- rbind(pgr_synch, expected_synch)

# ggplot(pgr_synch, aes(x=L2, y=value, fill=L2))+
#   geom_bar(stat="identity", position=position_dodge(0.9), color="grey")+
#   facet_wrap("L1", scales="free_x")+
#   scale_fill_manual(values = c("grey45", "steelblue", "slateblue4"))+
#   guides(fill=FALSE)+
#   scale_y_continuous(limits=c(0,1))+
#   xlab("Source of stochasticity")+
#   ylab("Species synchrony")

pgrsynch_plot <- ggplot(pgr_synch, aes(x=L2, y=value, color=L1, group=L1))+
  geom_line()+
  geom_point(size=5)+
  scale_color_manual(values = c("grey45", "steelblue", "slateblue4", "darkorange", "purple"),
                     name = "Site")+
  scale_y_continuous(limits=c(0,1))+
  xlab("Simulation")+
  ylab("Species synchrony")+
  theme_bw()

####
####  Compare observed and IPM simulated synchrony
####
nsites <- length(sites)

@@
